Exploratory Data Analysis: White Wine Quality by Andrew Kinnaird
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(gridExtra)
library(dplyr)
library(reshape2)
library(GGally)
library(scales)
library(memisc)
library(lattice)
library(MASS)
library(car)
library(reshape)
library(plyr)
library(psych)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data
setwd('/Users/thegatekeeper/Documents/Career/Udacity/Data Analyst Nanodegree/Exploratory Data Analysis/Project_Explore_and_Summarize_Data')
white_wine <- read.csv('wineQualityWhites.csv')
```

This report compares the physical characteristics and quality score of almost 
5,000 white wine samples.

# Univariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
dim(white_wine)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
head(white_wine)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
str(white_wine)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
sum(is.na(white_wine))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
summary(white_wine)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
white_wine$X <- NULL
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
white_wine$quality.factor <- factor(white_wine$quality, ordered = TRUE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
str(white_wine)
```

The data includes a quality score and 11 physcial characteristics of
4,898 white wine samples. There are no "NA" values. The variable "X" was 
removed. While the quality range allowed is 0 - 10, the actual range of quality 
is 3 - 9.

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(quality), data = white_wine) +
  geom_bar()
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
summary(white_wine$quality)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
count(white_wine$quality == 3)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
count(white_wine$quality == 9)
```

The quality distribution appears to be normally distributed from 3 - 9 with 
about 500 more quality scores of 5 than 7. And 15 more quality scores of 3 than
9.

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(fixed.acidity), data = white_wine) +
  geom_histogram(binwidth = .1)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
summary(white_wine$fixed.acidity)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(volatile.acidity), data = white_wine) +
  geom_histogram(binwidth = .01)
```
```{r echo=FALSE, message=FALSE, warning=FALSE, }
summary(white_wine$volatile.acidity)
```

Both fixed.acidity and volatile.acidity are right skewed each with extreme 
outliers. Do fixed.acidity and volatile.acidity correlate?

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(citric.acid), data = white_wine) +
  geom_histogram(binwidth = .01)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
summary(white_wine$citric.acid)
```

The variable citric.acid appears to be normally distributed with extreme 
outliers slightly below 1.25 and 1.75. Also, there is a peak near 0.5. Why is 
there a peak at 0.5?

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(residual.sugar), data = white_wine) +
  geom_histogram(binwidth = 0.1)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(residual.sugar), data = white_wine) +
  geom_histogram(binwidth = 0.1) +
  xlim(0,20)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(residual.sugar), data = white_wine) +
  geom_histogram(binwidth = .1) +
  scale_x_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  ggtitle("scale_x_log10")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
#Source: https://rpubs.com/mohammadshadan/meanmedianmode
invisible(sort(table(white_wine$residual.sugar)))
names(table(white_wine$residual.sugar))[table(white_wine$residual.sugar)==
                                          max(table(white_wine$residual.sugar))]
```


```{r echo=FALSE, message=FALSE, warning=FALSE, }
summary(white_wine$residual.sugar)
```

The majority of wines include about 1 gram/liter, with the mode being 1.2. 
Because of the right skew in the histogram, the data was transformed using log10
scale. This transformation revealed a bimodal distribution.

According the information provided with the data, "it's rare to find wines with 
less than 1 gram/liter [of sugar] and wines with greater than 45 grams/liter 
[of sugar] are considered sweet."

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(chlorides), data = white_wine) +
  geom_histogram(binwidth = .001)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(chlorides), data = white_wine) +
  geom_histogram(binwidth = .001) +
  xlim(0,0.1)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
summary(white_wine$chlorides)
```

With outliers removed, the chloride distribution appears normal.

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(free.sulfur.dioxide), data = white_wine) +
  geom_histogram(binwidth = 1)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(free.sulfur.dioxide), data = white_wine) +
  geom_histogram(binwidth = 1) +
  xlim(0,100)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
summary(white_wine$free.sulfur.dioxide)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(total.sulfur.dioxide), data = white_wine) +
  geom_histogram(binwidth = 1)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(total.sulfur.dioxide), data = white_wine) +
  geom_histogram(binwidth = 1) +
  xlim(0,300)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
summary(white_wine$total.sulfur.dioxide)
```

The variable "free.sulfur.dioxide" is a subset of "total.sulfur.dioxide," 
therefore total.sulfur.dioxide's mean and median are higher than 
free.sulfur.dioxide's. According to the information provided with the data, 
"at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and 
taste of wine." As SO2 levels increase and become evident, does the quality of 
the wine decrease?

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(density), data = white_wine) +
  geom_histogram(binwidth = .001)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(density), data = white_wine) +
  geom_histogram(binwidth = .001) +
  xlim(0.985,1.005)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
summary(white_wine$density)
```

Almost all density values are within a range of .005. With outliers removed, a 
slightly right skewed distribution is revealed. The distribution is skewed 
toward 1.000, the density of water (https://water.usgs.gov/edu/density.html). 

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(pH), data = white_wine) +
  geom_histogram(binwidth = 0.1)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
summary(white_wine$pH)
```

pH values are normally distributed. pH values increase in acidity as they 
approach 0 (https://water.usgs.gov/edu/ph.html). According to the information 
provided with the data, "most wines are between 3-4 on the pH scale."

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(sulphates), data = white_wine) +
  geom_histogram(binwidth = .01)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(sulphates), data = white_wine) +
  geom_histogram(binwidth = .01) +
  xlim(0.2, 0.9)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
summary(white_wine$sulphates)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
#Source: https://rpubs.com/mohammadshadan/meanmedianmode
invisible(sort(table(white_wine$sulphates)))
names(table(white_wine$sulphates))[table(white_wine$sulphates)==
                                          max(table(white_wine$sulphates))]
```

According to the information provided with the data, sulphates are "a wine 
additive which can contribute to sulfur dioxide gas (S02) levels, wich (sic) 
acts as an antimicrobial and antioxidant." Is this additive added at a 
standard ammount leading to the mode of 0.5?

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(alcohol), data = white_wine) +
  geom_histogram(binwidth = 0.1)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(alcohol), data = white_wine) +
  geom_histogram() +
  scale_x_log10() +
  ggtitle('scale_x_log10')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
#Source: https://rpubs.com/mohammadshadan/meanmedianmode
invisible(sort(table(white_wine$alcohol)))
names(table(white_wine$alcohol))[table(white_wine$alcohol)==
                                          max(table(white_wine$alcohol))]
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
summary(white_wine$alcohol)
```

Alcohol levels are skewed right with a mode of 9.4. In order to address the 
skew in the histogram, the alcohol data was transformed using a log10 
transoformation. Is there a correlation between increased alcohol and increased 
quality?

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(quality), data = subset(white_wine, residual.sugar < 1)) +
  geom_bar() +
  ggtitle("Quality of Wine with Low Residual Sugar") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
str(subset(white_wine$quality, white_wine$residual.sugar < 1))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
summary(subset(white_wine$quality, white_wine$residual.sugar < 1))
```

As stated earlier, "it's rare to find wines with less than 1 gram/liter and
wines with greater than 45 grams/liter are considered sweet." The subset to 
examines the quality of wines with less than 1 gram/liter of sugar. This subset 
of 77 white wines skews toward lower quality.

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(quality), data = subset(white_wine, residual.sugar >= 45)) +
  geom_bar(binwidth = 1) +
  scale_x_continuous(breaks = 1) +
  scale_y_continuous(breaks = 1) +
  xlim(3,9) +
  ylim(0,5) +
  ggtitle('Quality of "Sweet"" Wine') +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
summary(subset(white_wine$quality, white_wine$residual.sugar >= 45))
```

As stated earlier, "it's rare to find wines with less than 1 gram/liter and
wines with greater than 45 grams/liter are considered sweet." The subset to 
examines the quality of wines with more than 45 gram/liter of sugar. There is 
only one wine with 45 grams/liter of residual.sugar. It's quality score of 6 is 
above the entire sample size's mean of 5.878.

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(quality), data = subset(white_wine, residual.sugar >= 9.9)) +
  geom_bar() +
  ggtitle("Quality of Wine with High Residual Sugar") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
str(subset(white_wine$quality, white_wine$residual.sugar >= 9.9))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
summary(subset(white_wine$quality, white_wine$residual.sugar >= 9.9))
```

The subset examines the quality of wines with more than 9.9 gram/liter of 
sugar, which coincides with the third quartile of the entire white_wine sample. 
Similar to white wines with low amounts of sugar, this subset skews toward lower 
quality.

The mean quality for the entire sample of white wine is 5.878. The mean quality 
for wines with residual sugar of less than 1 is 5.377. The mean quality for 
wines with residual sugar greater than or equal to 45 is 6, but since there is 
only one wine with "high" residual sugar this result does not provide 
significant insight. The mean quality for wines with residual sugar greater than 
or equal to 9.9, the third quartile of the entire sample, is 5.739.

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(quality), data = subset(white_wine, alcohol >= 11.4)) +
  geom_bar() +
  ggtitle("Quality of Wine with High Alcohol Content") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
str(subset(white_wine$quality, white_wine$alcohol >= 11.4))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
summary(subset(white_wine$quality, white_wine$alcohol >= 11.4))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(quality), data = subset(white_wine, alcohol >= 9.50 & alcohol <= 11.40)) +
  geom_bar() +
  ggtitle("Quality of Wine with Interquartile Range Alcohol Content") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
str(subset(white_wine$quality, white_wine$alcohol >= 9.50 & 
             white_wine$alcohol <= 11.40))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
summary(subset(white_wine$quality, white_wine$alcohol >= 9.50 & 
                 white_wine$alcohol <= 11.40))
```

The mean quality for the entire sample of white wine is 5.878. The mean quality 
for wines with alcohol of greater than 11.4 is 6.423. The mean quality for 
wines with alcohol content within the interquartile range, i.e. 9.50 and 11.40, 
is 5.791.

# Univariate Analysis

### What is the structure of your dataset?
There are 4,898 white wines in the dataset with 12 features (fixed acidity, 
volatile acidity, citric acid, residual sugar, chlorides, free sulfur dioxide, 
total sulfur dioxide, density, pH, sulphates, alcohol, quality). Further 
information, including units, is below:

   - fixed acidity (tartaric acid - g / dm^3)
   - volatile acidity (acetic acid - g / dm^3)
   - citric acid (g / dm^3)
   - residual sugar (g / dm^3)
   - chlorides (sodium chloride - g / dm^3)
   - free sulfur dioxide (mg / dm^3)
   - total sulfur dioxide (mg / dm^3)
   - density (g / cm^3)
   - pH
   - sulphates (potassium sulphate - g / dm^3)
   - alcohol (% by volume)
   
   Output variable (based on sensory data): 
   - quality (score between 0 and 10)

Other observations:

The data set contains no 'NA' values

Quality is normally distributed. The range of quality is 3-9.

Most other variables are also normally distributed, with two notable exceptions:
residual sugar and alcohol.

Residual sugar's range is from 0.600 to 65.800 with a mean of 6.391 and mode of 
1.2, which created a right skewed histogram.

Alcohol's range is from 8 to 14.20 with a mean of 10.51 and mode of 9.4, which 
created a right skewed histogram.

### What is/are the main feature(s) of interest in your dataset?

The main feature of interest in the data set is quality. The analysis will seek 
to determine which features hihgly correlate with the quality of a white wine.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

While fixed acidity, volatile acidity, citric acid, residual sugar, chlorides, 
free sulfur dioxide, total sulfur dioxide, density, pH, sulphates, and alcohol
will likely all impact the quality of white wine, I suspect alcohol and residual 
sugar will impact the quality of white wine more than the other variables.

### Did you create any new variables from existing variables in the dataset?

No new variables were created.

### Of the features you investigated, were there any unusual distributions? \ 
Did you perform any operations on the data to tidy, adjust, or change the \ 
form of the data? If so, why did you do this?

The variable "X" was removed because the unique identifiers will not be 
utilized.

A log10 transformation was used on the right skewed residual.sugar and alcohol 
distributions. The resulting transformation of residual sugar appears bimoal 
with sugar content peaking just above 1 and just below 10. The resulting 
transformation of alcohol appears normal.

In several plots the x-axis was limited in order to remove outliers and 
establish a better understanding of the bulk of the data. Also, bin widths were 
adjusted to align with the significant figures for each variable.

# Bivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}
round(cor(subset(white_wine, select=-c(quality.factor))), 3)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
round(cor((subset(white_wine, select=-c(quality.factor))), white_wine$quality), 3)
```

The correlation betwwen all variables and 'quality'.

```{r echo=FALSE, message=FALSE, warning=FALSE, }
round(cor((subset(white_wine, select=-c(quality.factor))), white_wine$chlorides), 3)
```

The correlation betwwen all variables and 'chlorides'.

```{r echo=FALSE, message=FALSE, warning=FALSE, }
round(cor((subset(white_wine, select=-c(quality.factor))), white_wine$density), 3)
```

The correlation between all variables and 'density'.

```{r echo=FALSE, message=FALSE, warning=FALSE, }
round(cor((subset(white_wine, select=-c(quality.factor))), white_wine$alcohol), 3)
```

The correlation betwwen all variables and 'alcohol'.

```{r}
# http://www.sthda.com/english/wiki/scatter-plot-matrices-r-base-graphs
pairs.panels(subset(white_wine, select=-c(quality.factor)))
```

```{r}
# http://www.sthda.com/english/wiki/scatter-plot-matrices-r-base-graphs
pairs.panels(subset(white_wine, select=c(residual.sugar, chlorides, 
                                         total.sulfur.dioxide, density, alcohol,
                                         quality)))
```


Quality is the variable of focus. Therefore, correlations with quality are 
highlighted. The strongest correlations exist between quality and the following: 
- chlorides (-0.210)
- density (-0.307)
- alcohol (0.436)

Because alcohol has the strongest correlation to quality, the relationship 
between alcohol and all other variables is considered. The strongest 
correlations exist between alcohol and the following:
- density (-0.780)
- residual.sugar (-0.451)
- total sulfur dioxide (-0.449)

Between all variables, the strongest correlations exist between the following: 
- density and residual sugar (0.839)
- density and alcohol (-0.780)
- density and total sulfur dioxide  (0.530)

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(factor(quality), chlorides), data = white_wine) +
  geom_jitter(alpha = 0.05, color = 'deeppink') +
  geom_boxplot(alpha = 0.05, color = 'deeppink4') +
  stat_summary(fun.y = 'mean', geom = 'point', color = 'deeppink2') +
  ylim(0,.15) +
  labs(x = 'quality')
```

As quality increases, the concentration of chlorides decreases.

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(factor(quality), density), data = white_wine) +
  geom_jitter(alpha = 0.05, color = 'magenta') +
  geom_boxplot(alpha = 0.05, color = 'magenta4') +
  stat_summary(fun.y = 'mean', geom = 'point', color = 'magenta2') +
  ylim(0.995,1.0025) +
  labs(x = 'quality')
```

Acccording to the boxplot above, no inference can be made about the correlation 
between quality and density. 

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(factor(quality), alcohol), data = white_wine) +
  geom_jitter(alpha = 0.05, color = 'darkorchid') +
  geom_boxplot(alpha = 0.05, color = 'darkorchid4') +
  stat_summary(fun.y = 'mean', geom = 'point', color = 'darkorchid2') +
  ylim(8,14) +
  labs(x = 'quality')
```

Overall, quality increases as alcohol concentration increases. There is a 
decrease in alcohol concentration as quality increases from 3 to 5.

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(alcohol, density), data = white_wine) +
  geom_point(alpha = 0.1, color = 'cadetblue3') +
  geom_smooth(color = 'cadetblue4')
```

As the concentration of alcohol increases, density decreases.

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(alcohol, residual.sugar), data = white_wine) +
  geom_point(alpha = 0.1, color = 'blue') +
  geom_smooth(color = 'blue4')
```

As the concentration of alcohol increases, residual sugar decreases especially 
from an alcohol concentration of 8-10%.

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(alcohol, total.sulfur.dioxide), data = white_wine) +
  geom_point(alpha = 0.1, color = 'cyan') +
  geom_smooth(color = 'cyan4')
```

As the concentration of alcohol increases, total sulfur dioxide decreases.

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(density, residual.sugar), data = white_wine) +
  geom_point(alpha = 0.1, color = 'orange') +
  geom_smooth(color = 'orange4')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(density, residual.sugar), data = white_wine) +
  geom_point(alpha = 0.1, color = 'orange') +
  geom_smooth(color = 'orange4') +
  xlim(0.985,1.005) +
  ylim(0,30)
```

As density increases, residual sugar increases.

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(density, alcohol), data = white_wine) +
  geom_point(alpha = 0.1, color = 'darkgoldenrod') +
  geom_smooth(color = 'darkgoldenrod4')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(density, alcohol), data = white_wine) +
  geom_point(alpha = 0.1, color = 'darkgoldenrod') +
  geom_smooth(color = 'darkgoldenrod4') +
  xlim(0.985,1.015)
```

As density increases, alcohol concentration decreases.

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(density, total.sulfur.dioxide), data = white_wine) +
  geom_point(alpha = 0.1, color = 'chocolate') +
  geom_smooth(color = 'chocolate4')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(density, total.sulfur.dioxide), data = white_wine) +
  geom_point(alpha = 0.1, color = 'chocolate') +
  geom_smooth(color = 'chocolate4') +
  xlim(0.985,1.015)
```

As density increases, total sulfur dioxide increases.

```{r echo=FALSE, message=FALSE, warning=FALSE, }
print(subset(white_wine[c('residual.sugar','density','alcohol',
                          'total.sulfur.dioxide')], 
             white_wine$residual.sugar == 65.8))
```

In the plots above, there is an outlier at a density just below 1.04. This is a 
result of the high concentration of residual sugar in that wine.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features 
in the dataset?

Quality correlates most strongly with chlorides, density, and alcohol.

As quality increases, the concentration of chlorides decreases.

As quality increases, density decreases.

As quality increases, the concentration of alcohol increases. Except from a 
quality change from 3 - 5 where alcohol concentration decrease as quality 
increases.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

Because alcohol has the strongest correlation to quality, the relationship 
between alcohol and all other variables was considered. The strongest 
correlations exist between alcohol and the following: density, residual sugar, 
total sulfur dioxide.

As the concentration of alcohol increases, density decreases. 
"Alcohol, or ethanol, is the intoxicating agent found in beer, wine and liquor."
https://www.drugs.com/alcohol.html The density of ethanol is 0.7893 
(https://pubchem.ncbi.nlm.nih.gov/compound/ethanol#section=Density). 
Therefore, as the concentration of alcohol increases, the density of the wine 
decreases.

As the concentration of alcohol increases, residual sugar decreases especially 
from an alcohol concentration of 8-10%. This is a by product of alcohol 
production. "[W]hen winemaking happens, yeast eats sugar and makes ethanol 
(alcohol) as a by-product." (https://winefolly.com/review/sugar-in-wine-chart/) 
The sugar leftover after this process is called "residual sugar." 
(https://bit.ly/2K2nFv1) Therfore as the concentration of alcohol increases, the 
amount of sugar "eaten" by yeast increases, thus the amount residual sugar in 
the wine decreases and the percentage of alcohol increases simultaneously.

As the concentration of alcohol increases, total sulfur dioxide decreases. 
"Sulfur dioxide (SO2) is important in the winemaking process as it aids in 
preventing microbial growth[...]." 
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3472855/ "[Alcohol] acts 
synergistically and enhances the bacteria-killing effect of molecular SO2 
[sulfur dioxide], so high-alcohol wines require less SO2 protection" 
(https://www.extension.purdue.edu/extmedia/fs/fs-52-w.pdf). Therefore, as 
alcohol concentration increases, the amount of total sulfur dioxide decreases.

### What was the strongest relationship you found?

The strongest correlations between any variables were condsidered. The 
variables were density and each of the following: residual sugar, alcohol, and 
total sulfur dioxide.

As density increases, residual sugar increases. This correlation occurs becuase 
"[t]he more sugar that’s mixed into a measured amount of water, the higher the 
density of the mixture." 
(https://www.stevespanglerscience.com/lab/experiments/sugar-rainbow/)

As density increases, alcohol concentration decreases. "Alcohol, or ethanol, is 
the intoxicating agent found in beer, wine and liquor." 
(https://www.drugs.com/alcohol.html) The density of ethanol is 0.7893 
(https://pubchem.ncbi.nlm.nih.gov/compound/ethanol#section=Density). 
Therefore, as the concentration of alcohol increases, the density of the wine 
increases.

As density increases, total sulfur dioxide increases. The density of sulfur 
dioxide is 1.434. 
(https://pubchem.ncbi.nlm.nih.gov/compound/sulfur_dioxide#section=Density) 
Therefore as the concentration of sulfur dioxide increases, the density of wine 
increases.

# Multivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}
ggplot(aes(x = alcohol, y = density, color = quality.factor), data = white_wine) +
  geom_jitter() +
  scale_color_brewer(type = "seq")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(x = alcohol, y = density, color = quality.factor), data = white_wine) +
  geom_jitter() +
  ylim(0.985, 1.015) +
  scale_color_brewer(type = "seq")
```

As density decreases and alcohol increases, the quality scores appear to 
increase.

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(x = alcohol, y = residual.sugar, color = quality.factor), data = white_wine) +
  geom_jitter() +
  scale_color_brewer(type = "seq")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(x = alcohol, y = residual.sugar, color = quality.factor), data = white_wine) +
  geom_jitter() +
  ylim(0,35) +
  scale_color_brewer(type = "seq")
```

As residual sugar decreases and alcohol increases, the quality scores appear to  
increase.

```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(x = alcohol, y = total.sulfur.dioxide, color = quality.factor), 
       data = white_wine) +
  geom_jitter() +
  scale_color_brewer(type = "seq")
```

As total sulfur dioxide decrease and alcohol increase, quailty scores appear to 
increase.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

The relationship between density, alcohol, and quality appears to be the 
strongest. The relationships between residual sugar / alcohol appeared to show 
as residual sugar decreased and alcohol increased quality increased, but there 
does not appear to be a strong relationship. A similar result was shown by the 
plot of total sulfur dioxide and alcohol.

### Were there any interesting or surprising interactions between features?

As density and alcohol were most closely correlated with quality and density and 
alcohol were closely correlated with one another, the strength of the 
relationship betwen density, alcohol, and quality was not surprising.

------

# Final Plots and Summary

### Plot One

```{r}
print(mean(white_wine$quality))
```


```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}
ggplot(aes(quality.factor, fill = quality.factor), data = white_wine) +
  geom_bar() +
  scale_fill_manual(values=c("3" = "red", "4" = "orange", "5" = "yellow", "6" = 
                               "green", "7" = "blue", "8" = "violet", "9" = 
                               "purple"), guide=FALSE) +
  geom_vline(xintercept = (mean(white_wine$quality))-2) +
  annotate("text", x=(mean(white_wine$quality))-3.2, y=2000, 
           label=paste("Mean Score = ", round((mean(white_wine$quality)),3))) +
  ggtitle("Quality of Wine") +
  labs(x = "Score (1-10)", y = "Count") +
  theme(plot.title = element_text(hjust = 0.5))
```

### Description One

The plot "Quality of Wine" contains the quality score data ranging from 1-10 on 
4,898 wines. Wine quality is the feature of focus for this study. While the 
allowed range of scores is 1-10, only scores of 3-9 were given. The mean quality 
score is 5.878.

### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE, }
ggplot(aes(density, residual.sugar), data = white_wine) +
  geom_point(alpha = 0.1, color = 'orange') +
  geom_smooth(color = 'orange4') +
  ggtitle("The Correlation between Residual Sugar and Density") +
  labs(x = "Density (g / cm^3)", y = "Residual Sugar (g / dm^3)") +
  theme(plot.title = element_text(hjust = 0.5))
```

### Description Two

Plot two highlights the strongest correlation between two variables: 0.84 
between residual sugar and density. This relatsionship is explained by 
the density of sugar and the contents of wine. "A wine typically 
contains ethanol (~13%) [and] water (85%)... ." (https://bit.ly/2MXqAXh) The 
density of water is 1.00. (https://water.usgs.gov/edu/density.html) Because wine 
is 85% water, the densities of the wines in the sample are near 1.00. An 
increase of residual sugar (density of 1.56) increases the density of the wine
(https://bit.ly/2KNqlkd). This correlation lead to further research showing 
residual sugar, total sulfur dioxide, and alcohol are dependent upon one 
another. The amount of alcohol in a wine depends upon the amount of sugar 
"eaten" by yeast (https://winefolly.com/review/sugar-in-wine-chart/): the more 
sugar eaten, the more alcohol in the wine, the less residual sugar remains. 
(https://bit.ly/2K2nFv1) Then as the concentration of alcohol increases, total 
sulfur dioxide decreases, because less sulfur dioxide needs to be added to 
prevent microbial growth.(https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3472855/)
Therefore, a wine with less residual sugar will have more alcohol and a wine 
with more alcohol will have less total sulfur dioxide, creating a less dense 
wine.

### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}
ggplot(aes(x = alcohol, y = density, color = quality.factor), data = white_wine) +
  geom_jitter() +
  ylim(0.985, 1.005) +
  scale_color_brewer(type = "seq") +
  ggtitle("Quality of Wine Plotted on Alcohol and Density") +
  labs(x = "Alcohol Concentration (%)", y = "Density (g / cm^3)") +
  theme(plot.title = element_text(hjust = 0.5))
```

### Description Three

Plot three highlights the relationship between alcohol, density, and quality. As 
described in plot two, density and alcohol correlate with one another. These 
factors also appear to correlate with a higher quality score, as density 
decreases and alcohol increases (as described in plot two, residual sugar and 
total sulfur dioxide also decrease), the quality scores increase.
------

# Reflection

The analysis performed examined a sample of white wine which included 4,898 
observations of 13 variables. The variable of focus was the wine's quality 
score. Initial analysis sought to understand the relationship between 
quality and all other variables. Unfortunately, there wasn't a single variable 
which strongly correlated with quality. Surprisingly, this analysis revealed a 
stronger relationship between density and three other variables: residual sugar, 
total sulfur dioxide, and alcohol. Further research revealed the dependent 
nature of the relatinship of these variables. Quality was then plotted against 
alcohol and density which showed that quality does correlate with an increase in 
alcohol concentration and a decrease in density. Further research should seek to 
isolate the dependent factors within the dataset in order to control for amounts 
of residual sugar and total sulfur dioxide.